---
title: "Week2_Prj1"
author: "Ramya Sridharan"
date: "1/23/2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Getting Data - reading the given csv file into R

```{r}
data_full <- read.csv("activity.csv", header=TRUE)
```

Cleaning Data and looking at histogram -grouping data by steps and date and looking at the histogram

```{r}

data_full$date <- as.POSIXct(data_full$date, format="%Y-%m-%d")

data_full <- data.frame(date=data_full$date, 
                        weekday=tolower(weekdays(data_full$date)), 
                        steps=data_full$steps, 
                        interval=data_full$interval)
data_full_final <- cbind(data_full,daytype=ifelse(data_full$weekday == "saturday" | data_full$weekday == "sunday", "weekend", "weekday"))

sum_of_steps <- aggregate(steps~date,data_full_final, FUN = sum, na.rm=TRUE)
hist(sum_of_steps$steps, xlab = "Sum of Steps", ylab= "Frequency")
```

Looking at the mean and median of steps taken each day

```{r}
mean_steps <- mean(sum_of_steps$steps)
median_steps <- median(sum_of_steps$steps,na.rm=TRUE)
mean_steps
median_steps
```


Time series plot of the average number of steps taken - aggregate or group data by steps and interval and look at the mean. Followed bvy plotting of the time serie data.

```{r}
library(ggplot2)
avg_steps <- aggregate(steps~interval, data_full_final, mean)
ggplot(data=avg_steps,aes(interval,steps))+geom_line()+
  xlab("5 min interval")+
  ylab("Avg number of steps")
```

The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_5min <- avg_steps[which.max(avg_steps$steps),]
max_5min
```

Code to describe and show a strategy for imputing missing data

I looked at the total number of missing values in the dataset:

```{r}
missing <- is.na(data_full_final)
sum(missing)
sum(missing)/nrow(data_full_final)
```

That's about 13% of the data. So instead of ignoring the missing values, I am going to take the mean of the avg_steps and replace all NA values with this average.Before doing this, I am copying the data to another variable called data_imputed.

```{r}
data_imputed <- data_full_final

for (i in 1:nrow(data_imputed)){
  if (is.na(data_imputed[i,3])==TRUE){
    data_imputed[i,3]=mean(avg_steps$steps)}}

```

Let's look at NA count in data_imputed.

```{r}
sum(is.na(data_imputed))
```

Let's now look at the summary of data_full and data_imputed.

```{r}
summary(data_full_final)
summary(data_imputed)
```

We can see that there is no major change in the summary of the two data frames. 

Now let's plot the histogram of the steps and its mean and median values. 

```{r}
sum_of_steps_imputed <- aggregate(steps~date,data_imputed, FUN = sum)
hist(sum_of_steps_imputed$steps, xlab = "Sum of Steps", yalb = "Frequency")

mean_steps_im <- mean(sum_of_steps_imputed$steps)

median_steps_im <- median(sum_of_steps_imputed$steps,na.rm=TRUE)
mean_steps_im
median_steps_im
```

If we compare these to our previous histogram and mean/median values - there is no major impact due to imputing the NA values. 

Are there differences in activity patterns between weekdays and weekends?

```{r, echo=FALSE}
data_full$date <- as.POSIXct(data_full$date, format="%Y-%m-%d")

data_full <- data.frame(date=data_full$date, 
                        weekday=tolower(weekdays(data_full$date)), 
                        steps=data_full$steps, 
                        interval=data_full$interval)
data_full_final <- cbind(data_full,daytype=ifelse(data_full$weekday == "saturday" | data_full$weekday == "sunday", "weekend", "weekday"))
```

```{r}
head(data_imputed) #after replacing NA values with mean
```

Panel plot with time series plot:

```{r}
library(lattice)
mean_data <- aggregate(data_imputed$steps, 
                       by=list(data_imputed$daytype, 
                               data_imputed$weekday, data_imputed$interval), mean)
names(mean_data) <- c("daytype", "weekday", "interval", "mean")
head(mean_data)

xyplot(mean ~ interval | daytype, mean_data, 
       type="l", 
       lwd=0.25, col="red",
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```

